The Markov Link Method: a nonparametric approach to combine observations from multiple experiments

### 2018-08-07 paper/paper.tex
credible/confidence typo

\figgy{
\hfill{}\figgygrs{.4}{../images/simulationI}{(a) Estimation error, \\ varying number of samples \\ and number of sampling procedures.}\hfill{}
\hfill{}\figgygrs{.5}{../images/simulationIconf}{(b) Confidence intervals, \\ varying number of samples \\ and number of sampling procedures.}\hfill{}

@@

\caption{{\bf When can the MLM accurately estimate the link?} \label{fig:simI}  We use simulations to test the MLM.  In subfigure (a) we look at the error of the MLM link estimator in many different situations, varying the numbers of samples, sampling procedures, and the link parameters.  In each trial our `ground truth' link parameters picked uniformly at random from the parameter space.  The MLM estimator accurately recovers the ground truth from simulated data, as long as we have enough samples \emph{and} enough distinct sampling procedures.  In subfigure (b) we consider the MLM's credible intervals.  For a selection of trials we choose a random value of of $x,y$ and then compare the true link parameter $q^*(y|x)$ with the MLM confidence interval $C_{x,y}$.  We show the interval (in black) centered around the ground truth (in red).  Since the link parameters lie in $[0,1]$, the interval is always contained within $\pm1$ of the ground truth (this larger region is shown in gray).  The intervals mostly contain the truth, and get narrower with more samples as long as there are enough distinct sampling procedures.   In subfigure (c) we focus on the case of exactly $4$ sampling procedures.  In this case it is only possible to get very accurate estimates if the link takes on a particular form.  If the link is independent (i.e.\ $X,Y$ are independent) or if the link is invertible and deterministic (e.g.\ $X=Y$) then we can determine what the link is.  However, if the truth is more complex, small numbers of sampling procedures may make it impossible to determine the true link; this issue is discussed in detail in Section \ref{sec:casestudies}.}
}

@@

    The results are summarized in Figure \ref{fig:simI}.  In trials with more samples, the estimator generally has lower error.  However, to make the error actually converge to zero we need at least six distinct sampling procedures.  This figure also shows that the confidence intervals work correctly regardless of of the number of samples or sampling procedures; they include the ground truth with high probability.   With many samples and sampling procedures, the intervals are small and concentrated around the truth.  With fewer samples or sampling procedures, the intervals are sometimes forced to be larger.  However, our uncertainty is not the same for every parameter.  In some cases we are able to obtain a very tight credible interval for one particular parameter even though the overall estimator error is high.

@@

\figgy{
\hfill{}\figgygr{0.9}{../images/allenout}\hfill{}
\caption{{\bf MLM estimation of the link, with confidence interval.}  The central plot shows a portion of the MLM estimator applied to the Allen Institute data.  We look at a subselection of Standard types ($x$) and all of the Patch types ($y$).  For each combination $x,y$ we draw a rectangle whose color indicates the value of the estimator $\hat q(y|x)$.  We also show our confidence about these estimators.  On the left we indicate the lower bounds indicated by the credible interval, on the right we indicate upper bounds.  For some parameters our intervals are much tighter than others.  For example, it appears we have high confidence the Standard type `Vip Rspo4' is highly associated with the Patch type `Vip.'  By constrast, we have almost no idea what is associated with the Standard type `Meis2.'  The upper confidence interval bounds suggest that $q^*(y|\mathrm{Meis2})$ could be nearly 1 for many different Patch types.  Obviously it cannot be 1 for all of those types simultaneously, since $\sum_y q^*(y|\mathrm{Meis2})=1$, but the data simply doesn't tell us which $y$ carries the mass.  \label{fig:allenboot}}
}

We applied the MLM to this data to obtain point estimates $\tilde q(y|x)$ and confidence intervals $C_{x y}$.  In Figure \ref{fig:allenboot} we visualize these objects for selected values of $x,y$.  Each parameter has a slightly different story.  For example, for the Standard type $x=$`Vip Rspo4' and the Patch type $y=$`Vip,' we have that $C_{x,y}=[.92,1.0]$.  This supports the idea that the true link satisfies $q^*(y|x)\geq .92$: at least 92\% of the cells classified as type `Vip Rspo4' by the Standard method will be classified as `Vip' by the Patch method.  However, for other types there is more ambiguity.  For the Standard type $x=$`Lamp5 Egln 1' and the Patch types $y=$`Lamp' we have $C_{x y}=[0,1.0]$.  The data we have does not give us a definitive answer as to whether cells with Standard type `Lamp5 Egln 1' are being classified with Patch type `Lamp.'

The variability in the confidence regions suggests what we need to do in order to more closely determine the value of the calibration.  For example, if we could develop more unique sampling strategies which will include `Lamp5 Egln 1' cells, this would help us resolve our ambiguity about this aspect of the link.  Indeed, going back to the original data, it is easy to see why this ambiguity appeared in the first place.  Cells of the `Lamp5 Egln 1' type only appear in any number when using the sampling strategies `Gad2-IRES-Cre' and 'Slc32a1-IRES-Cre.'  Both of those sampling strategies yield a fairly similar mix of types when measured with Patch.  To get a better resolution of the calibration, we would need a sampling strategy that included `Lamp5 Egln 1' cells but represents a significantly different slice of the overall population.  For a particular proposed experiment, simulations such as those found in Section \ref{sec:simulations} can be used to determine how many samples might be required to get an accurate estimate of the link, $q^*$.

@@

The MLM appears to provide reasonable values in simulation and gives useful insight to real data.  Applied to neuronal cell-type data, the MLM clarified how two different cell-type classification systems are related.  We saw that some aspects of the two systems seem well-calibrated, but others we are less sure about.  The nature of the variability in the confidence intervals suggested directions for experiments which would further refine our understanding.

In future work, it would be interesting to use additional assumptions to help us estimate calibrations.  For example, as it stands the Markov Link Method will only yield narrow confidence intervals if each measurement tool returns one of a finite number of measurement outcomes.  In some cases, we may believe that similar measurement values should have similar probabilities.  Such smoothness assumptions would make it possible to apply the MLM to measurement tools which can return a continuum of values.

>>>>

\figgy{
\hfill{}\figgygrs{.4}{../images/simulationI}{(a) Estimation error, \\ varying number of samples \\ and number of sampling procedures.}\hfill{}
\hfill{}\figgygrs{.5}{../images/simulationIconf}{(b) Credible intervals, \\ varying number of samples \\ and number of sampling procedures.}\hfill{}

@@

\caption{{\bf When can the MLM accurately estimate the link?} \label{fig:simI}  We use simulations to test the MLM.  In subfigure (a) we look at the error of the MLM link estimator in many different situations, varying the numbers of samples, sampling procedures, and the link parameters.  In each trial our `ground truth' link parameters picked uniformly at random from the parameter space.  The MLM estimator accurately recovers the ground truth from simulated data, as long as we have enough samples \emph{and} enough distinct sampling procedures.  In subfigure (b) we consider the MLM's credible intervals.  For a selection of trials we choose a random value of of $x,y$ and then compare the true link parameter $q^*(y|x)$ with the MLM interval $C_{x,y}$.  We show the interval (in black) centered around the ground truth (in red).  Since the link parameters lie in $[0,1]$, the interval is always contained within $\pm1$ of the ground truth (this larger region is shown in gray).  The intervals mostly contain the truth, and get narrower with more samples as long as there are enough distinct sampling procedures.   In subfigure (c) we focus on the case of exactly $4$ sampling procedures.  In this case it is only possible to get very accurate estimates if the link takes on a particular form.  If the link is independent (i.e.\ $X,Y$ are independent) or if the link is invertible and deterministic (e.g.\ $X=Y$) then we can determine what the link is.  However, if the truth is more complex, small numbers of sampling procedures may make it impossible to determine the true link; this issue is discussed in detail in Section \ref{sec:casestudies}.}
}

@@

    The results are summarized in Figure \ref{fig:simI}.  In trials with more samples, the estimator generally has lower error.  However, to make the error actually converge to zero we need at least six distinct sampling procedures.  This figure also shows that the intervals work correctly regardless of of the number of samples or sampling procedures; they include the ground truth with high probability.   With many samples and sampling procedures, the intervals are small and concentrated around the truth.  With fewer samples or sampling procedures, the intervals are sometimes forced to be larger.  However, our uncertainty is not the same for every parameter.  In some cases we are able to obtain a very tight credible interval for one particular parameter even though the overall estimator error is high.

@@

\figgy{
\hfill{}\figgygr{0.9}{../images/allenout}\hfill{}
\caption{{\bf MLM estimation of the link, with credible interval.}  The central plot shows a portion of the MLM estimator applied to the Allen Institute data.  We look at a subselection of Standard types ($x$) and all of the Patch types ($y$).  For each combination $x,y$ we draw a rectangle whose color indicates the value of the estimator $\hat q(y|x)$.  We also show our confidence about these estimators.  On the left we indicate the lower bounds indicated by the credible interval, on the right we indicate upper bounds.  For some parameters our intervals are much tighter than others.  For example, it appears we have high confidence the Standard type `Vip Rspo4' is highly associated with the Patch type `Vip.'  By constrast, we have almost no idea what is associated with the Standard type `Meis2.'  The upper credible interval bounds suggest that $q^*(y|\mathrm{Meis2})$ could be nearly 1 for many different Patch types.  Obviously it cannot be 1 for all of those types simultaneously, since $\sum_y q^*(y|\mathrm{Meis2})=1$, but the data simply doesn't tell us which $y$ carries the mass.  \label{fig:allenboot}}
}

We applied the MLM to this data to obtain point estimates $\tilde q(y|x)$ and credible intervals $C_{x y}$.  In Figure \ref{fig:allenboot} we visualize these objects for selected values of $x,y$.  Each parameter has a slightly different story.  For example, for the Standard type $x=$`Vip Rspo4' and the Patch type $y=$`Vip,' we have that $C_{x,y}=[.92,1.0]$.  This supports the idea that the true link satisfies $q^*(y|x)\geq .92$: at least 92\% of the cells classified as type `Vip Rspo4' by the Standard method will be classified as `Vip' by the Patch method.  However, for other types there is more ambiguity.  For the Standard type $x=$`Lamp5 Egln 1' and the Patch types $y=$`Lamp' we have $C_{x y}=[0,1.0]$.  The data we have does not give us a definitive answer as to whether cells with Standard type `Lamp5 Egln 1' are being classified with Patch type `Lamp.'

The variability in the credible regions suggests what we need to do in order to more closely determine the value of the calibration.  For example, if we could develop more unique sampling strategies which will include `Lamp5 Egln 1' cells, this would help us resolve our ambiguity about this aspect of the link.  Indeed, going back to the original data, it is easy to see why this ambiguity appeared in the first place.  Cells of the `Lamp5 Egln 1' type only appear in any number when using the sampling strategies `Gad2-IRES-Cre' and 'Slc32a1-IRES-Cre.'  Both of those sampling strategies yield a fairly similar mix of types when measured with Patch.  To get a better resolution of the calibration, we would need a sampling strategy that included `Lamp5 Egln 1' cells but represents a significantly different slice of the overall population.  For a particular proposed experiment, simulations such as those found in Section \ref{sec:simulations} can be used to determine how many samples might be required to get an accurate estimate of the link, $q^*$.

@@

The MLM appears to provide reasonable values in simulation and gives useful insight to real data.  Applied to neuronal cell-type data, the MLM clarified how two different cell-type classification systems are related.  We saw that some aspects of the two systems seem well-calibrated, but others we are less sure about.  The nature of the variability in the credible intervals suggested directions for experiments which would further refine our understanding.

In future work, it would be interesting to use additional assumptions to help us estimate calibrations.  For example, as it stands the Markov Link Method will only yield narrow credible intervals if each measurement tool returns one of a finite number of measurement outcomes.  In some cases, we may believe that similar measurement values should have similar probabilities.  Such smoothness assumptions would make it possible to apply the MLM to measurement tools which can return a continuum of values.
